### 2016-05-12 msmfs.tex
Final version submitted
Minor changes per Anna.

% Enter here some information needed to fill in the template Title to appear
% on the front pages (will be filled in via the \sdpfrontpage command)
\newcommand{\bigdoctitle}{Implementation of the Rau-Cornwell MSMFS algorithm \xspace}
% Title to go in the "Document Status Sheet" Document number
\newcommand{\docnr}{TCC-SDP-151123-2\xspace}
% Context
\newcommand{\context}{(SDP.PIP)}
% Revision
\newcommand{\revision}{3\xspace}
% Author(s)
\newcommand{\docauthor}{T.J.\ Cornwell\xspace}
% Lead author (goes in the footer)
\newcommand{\leadauthor}{T.J.\ Cornwell\xspace}
% Release date (YYYY-MM-DD)
\newcommand{\docudate}{\today\ \currenttime}
% Document classification
\newcommand{\classification}{Unrestricted}
% Status of the document (draft/final/etc.)
\newcommand{\docstatus}{Draft\xspace}

@@

% Table with version numbers
\newcommand{\versiontable}{
  \begin{tabularx}{\textwidth}{|X|X|X|X|}
    \hline
    \bf{Version} & {\bf Date of issue} & {\bf Prepared by} & {\bf Comments}\\
    \hline
    0.3 & 27/04/2016 & T.J. Cornwell & Ready for review by Anna and Steve \\
    \hline
  \end{tabularx}
}

@@

\begin{enumerate}
\item Perform simulations to explore the need for MSMFS in SKA. Currently an L1 requirement directs SDP to implement MSMFS. However there is no justification given for this very computationally demanding processing. Indeed the promised excellent UV coverage of the SKA telescopes makes it at least questionable, although arguing from the work of Urvashi [RD??] does indicate that with MFS, the dynamic range would be limited to $10^4$ for wide frequency span. Either CASA or ASKAPsoft could be used for the reconstructions.
\item Adding to the desire to see some more justification of the requirement, we note that the L1 requirements do not address the number of scales required, the number of Taylor terms, or the possibility of sub-banding (all of which appear in the SDP L2 requirements. We note three possible outcomes of the simulations: MFS can be ignored, MFS is required and MSMFS is sufficient, or MFS is needed and MFMFS is not sufficient. In the case of the latter outcome, investigation of other deconvolution techniques (e.g. [RD14]) would be required.
\item Assess the importance of the considerations in Section \ref{sec:other}.
\end{enumerate}

@@

Urvashi has discussed the use of MSMFS for the JVLA (section 8.4.2.1 of [RD??]. With regard to just ignoring the effect of spectral index variations, she says:

@@

In the Multi-Scale CLEAN [RD11], the sky is modelled as a set of blobs. Performing this decomposition using a gradient search algorithm is very demanding of resources. Instead, in MSCLEAN, a greedy algorithm obtained by generalising the CLEAN algorithm is used. The greedy algorithm locates the amplitude, scale and location of the blob by a straightforward search in the set of residual images, each smoothed with the appropriate blob. For each minor cycle this produces an amplitude, location, and scale. The smoothed residual images and model images are updated, and iteration proceeds.

@@

However, the Hessian is still useful since some of the diagonal terms can used to decouple the pixels in Taylor term space as well as possible. In practice, the Hessian (for each scale) for the peak of the PSF in Taylor term space is of modest size and can be inverted and used. The greedy algorithm then locates the scale and location in the reference-frequency image (as convolved with the set of scales). For each minor cycle this produces a location, scale, and values for all Taylor terms. The residual images and model images are updated, and iteration proceeds. The actual algorithm is shown in Algorithm \ref{algo:CASA} (CASA form) and Algorithm \ref{algo:ASKAP} (ASKAP form).

In summary, MSMFS decouples Taylor terms using the Hessian in Taylor-space evaluated at the peak of the PSF, and uses a greedy CLEAN-like algorithm to decouple in scale and sky position. The scales can be decoupled by an orthogonalisation process prior to deconvolution but the resulting images tend to have non-physical regions of negative brightness.

@@

\begin{description}
\item[Polarisation] Only Stokes I can be modelled using MSMFS. Modelling the Faraday Rotation is possible but there is no existing algorithm.
\item[Frequency] O(500) coarse channels are required to avoid bandwidth smearing/ There is no advantage to finer channels (see e.g. section 8.4 in [RD??]).
\item[Sub-bands] O(1-3) partitions of the coarse channels are required to ensure that the power law approximation is adequate.
\item[Time] O(43200) ~ O(1s) correlator samples are required to avoid smearing on the longest baselines
\item[Facet] O(100) facets to limit the size of the AW projection kernel
%\item[Snapshot time] O(100s) long gridding/degridding sequences are required to limit smearing due to w-plane rotation
\item[Scales] O(5) scales are typically required to model extended structure (see section 8.4 in [RD??])
\item[Taylor terms] O(5) terms are required to represent the spectral behaviour for moderately complex spectrum sources (see section 8.4 in [RD??])
\end{description}

>>>>

% Enter here some information needed to fill in the template Title to appear
% on the front pages (will be filled in via the \sdpfrontpage command)
\newcommand{\bigdoctitle}{Implementation of the Rau-Cornwell MSMFS algorithm \xspace}
% Title to go in the "Document Status Sheet" Document number
\newcommand{\docnr}{TCC-SDP-151123-2\xspace}
% Context
\newcommand{\context}{(SDP.PIP)}
% Revision
\newcommand{\revision}{1.0\xspace}
% Author(s)
\newcommand{\docauthor}{T.J.\ Cornwell\xspace}
% Lead author (goes in the footer)
\newcommand{\leadauthor}{T.J.\ Cornwell\xspace}
% Release date (YYYY-MM-DD)
\newcommand{\docudate}{\today\ \currenttime}
% Document classification
\newcommand{\classification}{Unrestricted}
% Status of the document (draft/final/etc.)
\newcommand{\docstatus}{Released\xspace}

@@

% Table with version numbers
\newcommand{\versiontable}{
  \begin{tabularx}{\textwidth}{|X|X|X|X|}
    \hline
    \bf{Version} & {\bf Date of issue} & {\bf Prepared by} & {\bf Comments}\\
    \hline
    0.3 & 27/04/2016 & T.J. Cornwell & Ready for review by Anna and Steve \\
    1.0 & 12/05/2016 & T.J. Cornwell & To be signed \\
    \hline
  \end{tabularx}
}

@@

\begin{enumerate}
\item Perform simulations to explore the need for MSMFS in SKA. Currently an L1 requirement directs SDP to implement MSMFS. However there is no justification given for this very computationally demanding processing. Indeed the promised excellent UV coverage of the SKA telescopes makes it at least questionable, although arguing from the work of Urvashi [RD07] does indicate that with MFS, the dynamic range would be limited to $10^4$ for wide frequency span. Either CASA or ASKAPsoft could be used for the reconstructions.
\item Adding to the desire to see some more justification of the requirement, we note that the L1 requirements do not address the number of scales required, the number of Taylor terms, or the possibility of sub-banding (all of which appear in the SDP L2 requirements. We note three possible outcomes of the simulations: MFS can be ignored, MFS is required and MSMFS is sufficient, or MFS is needed and MFMFS is not sufficient. In the case of the latter outcome, investigation of other deconvolution techniques (e.g. [RD14]) would be required.
\item Assess the importance of the considerations in Section \ref{sec:other}.
\end{enumerate}

@@

Urvashi has discussed the use of MSMFS for the JVLA (section 8.4.2.1 of [RD07]. With regard to just ignoring the effect of spectral index variations, she says:

@@

In the Multi-Scale CLEAN [RD11], the sky is modelled as a set of blobs. Performing this decomposition using a gradient search algorithm is very demanding of resources. Instead, in MSCLEAN, a greedy algorithm obtained by generalising the CLEAN algorithm is used. The greedy algorithm locates the amplitude, scale and location of the blob by a straightforward search in the set of residual images, each smoothed with the appropriate blob. For each minor cycle this produces an amplitude, location, and scale. The smoothed residual images and model images are updated, and iteration proceeds\footnote{To exclude any errors in transcription, the LaTeX code for the equations in this section is drawn directly from [RD16] via arxiv.org}.

@@

However, the Hessian is still useful since some of the terms can used to decouple the pixels in Taylor term space as well as possible. In practice, the Hessian (for each scale) for the peak of the PSF in Taylor term space is of modest size and can be inverted and used. The greedy algorithm then locates the scale and location in the reference-frequency image (as convolved with the set of scales). For each minor cycle this produces a location, scale, and values for all Taylor terms. The residual images and model images are updated, and iteration proceeds. The actual algorithm is shown in Algorithm \ref{algo:CASA} (CASA form) and Algorithm \ref{algo:ASKAP} (ASKAP form).

In summary, MSMFS decouples Taylor terms using the (non-diagonal) Hessian in Taylor-space evaluated at the peak of the PSF, and uses a greedy CLEAN-like algorithm to decouple in scale and sky position. The scales can be decoupled by an orthogonalisation process prior to deconvolution but the resulting images tend to have non-physical regions of negative brightness.

@@

\begin{description}
\item[Polarisation] Only Stokes I can be modelled using MSMFS. Modelling the Faraday Rotation is possible but there is no existing algorithm.
\item[Frequency] O(500) coarse channels are required to avoid bandwidth smearing/ There is no advantage to finer channels (see e.g. section 8.4 in [RD07]).
\item[Sub-bands] O(1-3) partitions of the coarse channels are required to ensure that the power law approximation is adequate.
\item[Time] O(43200) ~ O(1s) correlator samples are required to avoid smearing on the longest baselines
\item[Facet] O(100) facets to limit the size of the AW projection kernel
%\item[Snapshot time] O(100s) long gridding/degridding sequences are required to limit smearing due to w-plane rotation
\item[Scales] O(5) scales are typically required to model extended structure (see section 8.4 in [RD07])
\item[Taylor terms] O(5) terms are required to represent the spectral behaviour for moderately complex spectrum sources (see section 8.4 in [RD07])
\end{description}
